### 2016-08-27 FrontBackmatter/Introduction.tex
Cosmetics of the notation
Massage the introduction

\chapter*{Introduction}
\addtocontents{toc}{\protect\vspace{\beforebibskip}} % to have the bib a bit from the rest in the toc
\addcontentsline{toc}{section}{\tocEntry{Introduction}}

@@

@@

\todo[inline]{CURRENTLY VERBATIM FROM MASTER THESIS}
A similar situation in encountered when quantum chemistry deals with the
description of complex systems, such as solutions or proteins. Unlike
relativity, the inclusion of the environment in the description of a
chemical system has been considered from the early stages in the
evolution of quantum chemistry.
Already in 1936 \citeauthor{Onsager1936-wf} had proposed a model to
account for the presence of a liquid on point-like electric
moments.~\autocite{Onsager1936-wf}

As the development of our knowledge of solutions reflects to some extent
the development of chemistry itself, this interest is very well
justified.
Despite the large number of reactions known to happen in the solid state
and in gas phase, we can state that almost always chemistry happens in
solution.~\autocite{Reichardt2010-le}

The theoretical treatment of environment effects suffers from a
\emph{dimensionality} problem, since even the most simplified picture of
the system under consideration would require taking into account
500--1000 atoms, at least. Application of the most accurate methods in
quantum chemistry is therefore impossible and not even desirable: as
always in problems with a high dimensionality, the microscopic detail in
the physical description cannot account for the macroscopic behaviour of
the system. This is an important point that will be discussed in greater
detail later.
Models must be devised to overcome the dimensionality ``disease'': the
ability to account, at least qualitatively, for environment effects is
rather fundamental in all experimental branches of
chemistry.~\autocite{Anderson1972-ai, Winsberg2010-sy, Kovac2011-ew}

@@

@@

We should state more precisely what we mean by solution, here and in the
following we will always refer to a system were the number of solvent
molecules exceeds by far the number of solute molecules. From an
experimental point of view, it is an \emph{extremely} dilute solution.
We will follow the expositions given by Tomasi in refs.~
\noparcite{Tomasi2004-dc} and \noparcite{Tomasi2007-es}.

Our observables of interest are the same as \emph{in vacuo}, \ie
relative energies, spectroscopic parameters and so forth. These
observables are influenced by the presence of the solvent and this is
what is commonly called \emph{solvent effect}.
Solvent effects are, customarily but conveniently, classified as:
\begin{itemize}
\item[.] \emph{direct}. The presence of the solvent modifies the solute
  electronic density, all observables depending on the density are
  accordingly perturbed;
\item[.] \emph{indirect}. In the Born--Oppenheimer picture the
  electronic energy acts as \ac{PES} for the
  nuclear motion. The most striking solvent effect is thus the
  modification of the equilibrium geometry of a molecular solute.
  Aminoacids are notable examples of this behaviour, exhibiting
  different predominant conformers in different solvents;
\item[.] \emph{local field}. The presence of the solvent can alter the
  response of the system to applied fields, locally modifying the
  applied field itself.~\autocite{Cammi1998-jp, Pipolo2014-sd}
\item[.] \emph{dynamic}. When considering electronic excitations we
  always work in a Franck--Condon framework. This is also true for
  excitation processes in solution. In addition, we will have to
  consider not only the geometric relaxation of the solute, but also the
  relaxation of the surrounding medium;
\item[.] \emph{specific}. A brief description of specific solvent
  effects is not possible, due to their extremely varied nature. The
  name specific comes from the fact that their existence depends on the
  particular solute-solvent pair, rather than only on one of the
  partners. It appears thus safe to say that all the effects that must
  be described considering the atomistic nature of the solute-solvent
  interaction are specific.
\end{itemize}

It is, again, customary and convenient to divide the models proposed in
two classes, according to the microscopic description of the solvent
these models give. Hence we distinguish between:
\begin{itemize}
 \item[.] \emph{discrete} (or \emph{explicit}). The degrees of freedom
   associated with the solvent molecules are dealt with explicitly as
   are those associated with the solute. The two sets of coordinates can
   be treated at different levels of theory;
 \item[.] \emph{continuum} (or \emph{implicit}). Only the degrees of
   freedom associated with the solute are dealt with explicitly, the
   solvent being replaced with a structureless continuum characterized
   by its bulk properties.
\end{itemize}

@@

\autocite{Anderson1972-ai}

This will be the subject of Chapter \ref{ch:CSM}

@@

>>>>

\chapter*{Introduction}
\addtocontents{toc}{\protect\vspace{\beforebibskip}} % to have the bib a bit from the rest in the toc
\addcontentsline{toc}{chapter}{\tocEntry{Introduction}}

\renewcommand{\thefigure}{\Alph{figure}}

@@

\epigraph{
[...] og fordi jeg alltid har hatt en dragning mot det skjulte og hemmelige.
    }{
    --- \textsc{K. O. Knausgård}, \textit{Om Høsten}}

\begin{figure}[tb]
\centering
\scalebox{0.8}{\input{gfx/RQC-axes.tikz}}
\caption{The three dimensions of Relativistic Quantum Chemistry. The computational scaling of a relativistic model chemistry is
$ S=\textcolor{red}{y}(\textcolor{blue}{x})f(\textcolor{blue}{x})\textcolor{green}{M}^{\textcolor{blue}{x}} $. The expense of relativity is in
the appearance of a new prefactor, not in the $M$-dependence of the particular method.}
\label{fig:RQC-axis}
\end{figure}
and look at different axes.

\section*{What Are the Electrons Really Doing in Molecules?}

\todo[inline]{Write something about the scope of molecular quantum
mechanics.}

This question was posed by R.~S.~Mulliken over a half-century ago, as
lecture title for the acceptance of the Gilbert N.~Lewis award in 1960.

@@

@@

Chemistry can be largely considered a \emph{wet} science: almost always
chemical phenomena happen in a liquid environment.~\autocite{Reichardt2010-le}
We hereby define a "solution", or more generally an "environment", as
a system were the number of solvent molecules exceeds by far the number
of solute molecules.~\autocite{Tomasi2004-dc, Tomasi2007-es}
It is then clear that theoretical and computational approaches to such a
problem will necessarily suffer from a \emph{dimensionality disease}.
The number of degrees of freedom to be taken into account is, in
principle, so large, that even the most advanced computing machines
would have a hard time computing the desired observables.
Moreover, on an interpretive level, it would not even be desirable to
have such a detailed insight.
As is well known from statistical mechanics, microscopic detail cannot
account for the macroscopic behaviour.~\autocite{Hill1960-ql,
Hansen2013-io}
To tame this complexity and cure the disease, one must devise
\emph{models} that simplify physical reality, while offering tools for
understanding reality and predicting new and exciting
phenomena.~\autocite{Anderson1972-ai, Winsberg2010-sy, Kovac2011-ew}
One of the earlier attempts at tackling the problem of solvation is due
to \citeauthor{Onsager1936-wf}. His was a rather crude model, but one
that has had a lasting impact and informs much of the developments that
will be presented in this thesis.~\autocite{Onsager1936-wf}

Before introducing our model of choice, let us consider how an
environment might affect molecular observables of interest.
Environment effects are usually classified as:
\begin{itemize}
\item \emph{direct}.
  These effect stem straightforwardly from the modification underwent by
  the solute electronic density when interacting with the environment.
\item \emph{Indirect}.
  It is common for solutes to exhibit different minimum-energy
  conformations in different environments. These effects are commonly
  labelled as indirect.
\item \emph{Local field}.
  Light-matter interactions are also affected by the environment. Local
  modifications of externally applied fields subtly influence molecular
  responses.~\autocite{Cammi1998-jp, Pipolo2014-sd}
\item \emph{Dynamic}.
  Probing the nature of molecular excited states has enormous
  technological impact. The presence of the environment radically
  influences excited states, since relaxation processes in the medium
  become important.
\item \emph{Specific}. This catch-all category includes all effects
  stemming from the peculiar solute-solvent pair interactions that
  cannot be fully described under any of the previous labels.
  In general, modelling such effects demands an atomistic level of
  detail.
\end{itemize}

Faced with the problem of describing such a diverse array of effects,
two main models have emerged in the past decades, each with its
strengths and weaknesses.
Both can be classified as \emph{multiscale} (or \emph{focused})
models~\autocite{Nobel2013} and hinge on the same idea: treat different
parts of the system with different methods and couple these methods by
bridging "scales" at the boundary, \emph{vide infra}.
While both models treat the molecular degrees of freedom at the quantum
mechanical level, their approach to the microscopic description of the
degrees of freedom of the environment differs:
\begin{itemize}
 \item
   \emph{Discrete} (or \emph{explicit}) models explicitly treat those
   degrees of freedom.
   This is either achieved by a cheaper quantum mechanical
   method~\autocite{Vreven2006-gx} or by \ac{MM}.~\autocite{Senn2009-sk}
   In the latter approach, commonly dubbed \acs{QM}/\acs{MM}, the \acs{MM}
   region can either be polarizable~\autocite{MMPol} or non
   polarizable.~\autocite{MM}
 \item
   \emph{Continuum} (or \emph{implicit}) completely remove the degrees
   of freedom of the environment from the picture, replacing it with a
   structureless continuum.
   The effect of this continuum is described, classically, \emph{via}
   its bulk properties.~\autocite{Onsager1936-wf, Miertus1981-mm}
\end{itemize}
\acs{QM}/\acs{MM} models can capture, albeit approximately, the effect
of the atomistic nature of the environment on the active part of the
system.
However, they demand a statistical average of environment configurations
to yield results of any significance. Moreover, a large number
\emph{cutoff} for the \acs{MM} region is usually required to converge
long-range electrostatic interactions.~\autocite{Something here??}
Continuum models avoid both problems at once. Statistical averaging is
built into the model \emph{via} their parametrization by
means of the environment's bulk properties, such as the permittivity.
In addition, long-range electrostatics is treated exactly.
Unfortunately, atomistic detail is lost and it is then impossible to
recover a satisfactory description of specific effects.
To partly alleviate these sources of error, the \acs{QM}/\acs{MM} and
\acs{QM}/Continuum methods can and have been successfully combined to
yield the three-layer \acs{QM}/\acs{MM}/Continuum method.~\autocite{Curutchet, Steindal Lipparini}
Figure \ref{fig:qm-to-multiscale} schematically portrays the transition
from a full \acs{QM} model of the relevant system to its multiscale
representations.

\begin{figure}[tb]
\missingfigure{Add picture with multiscale models}
\caption{From quantum mechanical models to multiscale models.}
\label{fig:qm-to-multiscale}
\end{figure}

Notice that we have deliberately ruled out so-called \emph{cluster}
models from the above discussion.~\autocite{ClusterModels}
These approaches replace the actual physical setting with a suitable
truncation of the whole solute+solvent system, the \emph{model system}
and treat it within a chosen quantum mechanical level of theory.
Cluster models can be used to benchmark more approximate multiscale
models, but their description is outside the scope of this thesis.

Chapter \ref{ch:CSM} will present an overview of the \ac{PCM} for
solvation.
I will present a nontechnical discussion of the mathematical details of
the model and an outline of current methodologies for the solution of
the associated governing equations.
Borrowing from the work of \citeauthor{Lipparini2010-be},\autocite{Lipparini2010-be,
Lipparini2015-lq} I will introduce a unifying theoretical formalism for
\acs{QM}/Continuum, \acs{QM}/\acs{MM} and \acs{QM}/\acs{MM}/Continuum models that will
be extensively used throughout the thesis.

@@

@@

@@

Characterizing, explaining and predicting the effect of the chemical
environment on a large number of measurable properties requires,
however, a synergistic experimental and theoretical approach.
